---
title: "BUSIE702_Finance_Project"
author: "Xiaoliang Zhang (1300019)"
date: 'July 11st, 2022'
output: pdf_document
---

```{r}
library(readxl)

# Reading the data
finance_df <- read_excel("Finance data for practicum.xlsx")
finance_df=finance_df[-505,]
```

# checking data status
```{r}
#glimpse(finance_df)

str(finance_df)

summary(finance_df)
```


## Change coloum name 
```{r}
names(finance_df)
colnames(finance_df) <- c("Equity", "Current.Market.Cap","Last.Price","Alpha","Beta",
                          "Revenue.Growth.Year.over.Year","ROA","ROE","Profit.Margin",
                          "Asset.Turnover","Financial.Leverage","EBIT.Interest","Price.Earnings.Ratio",
                          "Price.to.Sales.Ratio","Short.and.Long.Term.Debt","Total.Equity",
                          "Operating.Return.on.Total.Invested.Capital","Weighted.Average.Cost.of.Cap",
                          "Free.Cash.Flow","Price.to.Book.Ratio","Industry.Group","WACC.Economic.Value.Added",
                          "Total.Compensation.Paid.to.Executives")

# colnames(finance_df)[which(names(finance_df) == "...1")] <- "Equity"

```

## Change to numeric/factor 
```{r}
## change to numeric 
finance_df$Alpha <- as.numeric(finance_df$Alpha)

finance_df$Beta <- as.numeric(finance_df$Beta)

finance_df$ROA <- as.numeric(finance_df$ROA)

finance_df$ROE <- as.numeric(finance_df$ROE)

finance_df$Asset.Turnover<- as.numeric(finance_df$Asset.Turnover)

finance_df$Financial.Leverage<- as.numeric(finance_df$Financial.Leverage)

finance_df$EBIT.Interest<- as.numeric(finance_df$EBIT.Interest)

finance_df$Price.Earnings.Ratio<- as.numeric(finance_df$Price.Earnings.Ratio)

finance_df$Price.to.Sales.Ratio<- as.numeric(finance_df$Price.to.Sales.Ratio)

finance_df$Operating.Return.on.Total.Invested.Capital<- as.numeric(finance_df$Operating.Return.on.Total.Invested.Capital)

finance_df$Price.to.Book.Ratio<- as.numeric(finance_df$Price.to.Book.Ratio)

finance_df$WACC.Economic.Value.Added<- as.numeric(finance_df$WACC.Economic.Value.Added)

finance_df$Total.Compensation.Paid.to.Executives<- as.numeric(finance_df$Total.Compensation.Paid.to.Executives)

## change to factr 
finance_df$Equity <- as.factor(finance_df$Equity)

finance_df$Industry.Group <- as.factor(finance_df$Industry.Group)


summary(finance_df)
```

# checking missing data
```{r}
colSums(is.na(finance_df)) # check missing value 

barplot(colSums(is.na(finance_df))) # check missing value (#N/A N/A)
```


# Deal Missing data with 'mean'
```{r}
finance_df$Alpha[is.na(finance_df$Alpha)] <- 
  mean(finance_df$Alpha, na.rm = TRUE)

finance_df$ROA[is.na(finance_df$ROA)] <- 
  mean(finance_df$ROA, na.rm = TRUE)

finance_df$Asset.Turnover[is.na(finance_df$Asset.Turnover)] <- 
  mean(finance_df$Asset.Turnover, na.rm = TRUE)

finance_df$Price.Earnings.Ratio[is.na(finance_df$Price.Earnings.Ratio)] <- 
  mean(finance_df$Price.Earnings.Ratio, na.rm = TRUE)

finance_df$Beta[is.na(finance_df$Beta)] <- 
  mean(finance_df$Beta, na.rm = TRUE)

finance_df$ROE[is.na(finance_df$ROE)] <- 
  mean(finance_df$ROE, na.rm = TRUE)

finance_df$Financial.Leverage[is.na(finance_df$Financial.Leverage)] <- 
  mean(finance_df$Financial.Leverage, na.rm = TRUE)

finance_df$Price.to.Sales.Ratio[is.na(finance_df$Price.to.Sales.Ratio)] <- 
  mean(finance_df$Price.to.Sales.Ratio, na.rm = TRUE)

finance_df$Operating.Return.on.Total.Invested.Capital[is.na(finance_df$Operating.Return.on.Total.Invested.Capital)] <- 
  mean(finance_df$Operating.Return.on.Total.Invested.Capital, na.rm = TRUE)

finance_df$ Price.to.Book.Ratio[is.na(finance_df$ Price.to.Book.Ratio)] <- 
  mean(finance_df$ Price.to.Book.Ratio, na.rm = TRUE)

finance_df$Total.Compensation.Paid.to.Executives[is.na(finance_df$Total.Compensation.Paid.to.Executives)] <- 
  mean(finance_df$Total.Compensation.Paid.to.Executives, na.rm = TRUE)

finance_df$ EBIT.Interest[is.na(finance_df$ EBIT.Interest)] <- 
  mean(finance_df$ EBIT.Interest, na.rm = TRUE)

colSums(is.na(finance_df))
```

```{r}
library(dplyr)
library(car)

plot(finance_df$Current.Market.Cap)
hist(finance_df$Current.Market.Cap)

boxplot(finance_df$Current.Market.Cap)

boxplot.stats(finance_df$Current.Market.Cap)$out

#summary(finance_df$Current.Market.Cap)
```



# correlation (left)
```{r}
# pcs in R using prcomp()
library(factoextra)
library(ggplot2)
library(corrplot)
library(GGally)

numeric_df = finance_df
numeric_df$Equity <- as.numeric(as.factor(numeric_df$Equity))
numeric_df$Industry.Group <- as.numeric(as.factor(numeric_df$Industry.Group))

cor(numeric_df)

#corrplot(cor(numeric_df), method = 'color',type = 'lower') # no high correlation
ggcorr(numeric_df, method = c("everything", "pearson")) 
```

## group by 'Industry.Group'

```{r}
library(dplyr)
library(tidyverse)

# finance_df$Industry.Group = as.character(finance_df$Industry.Group)

## group by 'Industry.Group' keep mean 'Current.Market.Cap' & 'Free.Cash.Flow'
finance_gb <- finance_df %>% group_by(Industry.Group)
finance_gb <- finance_gb %>% group_by(Industry.Group) %>% 
  summarise (
Current.Market.Cap = mean(Current.Market.Cap),
  Free.Cash.Flow = mean(Free.Cash.Flow),
  Total.Equity = mean(Total.Equity),
  Last.Price = mean(Last.Price),
  WACC.Economic.Value.Added = mean(WACC.Economic.Value.Added),
  Total.Compensation.Paid.to.Executives = mean(Total.Compensation.Paid.to.Executives),
  ROA = mean(ROA),
  Short.and.Long.Term.Debt = mean(Short.and.Long.Term.Debt),
  Operating.Return.on.Total.Invested.Capital = mean(Operating.Return.on.Total.Invested.Capital),
  Short.and.Long.Term.Debt = mean(Short.and.Long.Term.Debt),
  Asset.Turnover = mean(Asset.Turnover),
  Price.to.Sales.Ratio = mean(Price.to.Sales.Ratio)
  )

#finance_gb %>% filter(Current.Market.Cap == max(Current.Market.Cap))
#plot(finance_gb)
 
```

## barplot (Current.Market.Cap)
```{r}
# load the library
library(forcats)

# Reorder following the value of another column:
finance_gb %>%
  mutate(Industry.Group = fct_reorder(Industry.Group, desc(Current.Market.Cap))) %>%
  ggplot( aes(x=Industry.Group, y=Current.Market.Cap)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") 
```

## Linear Scatter plot (Current.Market.Cap & Free.Cash.Flow) 
```{r}
# Library
library(ggplot2)
library(hrbrthemes)

p <- ggplot(finance_gb, aes(x=Free.Cash.Flow, y=Current.Market.Cap)) +
  geom_point(color=as.numeric(finance_gb$Industry.Group)) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
p
```


## Circular barplot (Current.Market.Cap & Free.Cash.Flow)
```{r}
library(dplyr)
library(ggplot2)
library(stringr)

plt <- ggplot(finance_gb) +
  # Make custom panel grid
  geom_hline(
    aes(yintercept = y), 
    data.frame(y = c(0:3) * 1000),
    color = "lightgrey"
  ) + 
  # Add bars to represent the cumulative track lengths
  # str_wrap(Industry.Group, 5) wraps the text so each line has at most 5 characters
  # (but it doesn't break long words!)
  geom_col(
    aes(
      x = reorder(str_wrap(Industry.Group, 5), Free.Cash.Flow),
      y = Free.Cash.Flow,
      fill = Free.Cash.Flow
    ),
    position = "dodge2",
    show.legend = TRUE,
    alpha = .9
  ) +
  
  # Add dots to represent the mean gain
  geom_point(
    aes(
      x = reorder(str_wrap(Industry.Group, 5), Free.Cash.Flow),
      y = Free.Cash.Flow
    ),
    size = 0.5,
    color = "gray12"
  ) +
   
  # Lollipop shaft for mean gain per region
  geom_segment(
    aes(
      x = reorder(str_wrap(Industry.Group, 5), Free.Cash.Flow),
      y = 0,
      xend = reorder(str_wrap(Industry.Group, 5), Free.Cash.Flow),
      yend = 3000
    ),
    linetype = "dashed",
    color = "gray12"
  ) + 
  
  # Make it circular!
  coord_polar()

plt
```


## Bubble chart (Current.Market.Cap & Free.Cash.Flow & Total.Equity) 
```{r}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

# Most basic bubble plot
finance_gb %>%
  arrange(desc(Current.Market.Cap)) %>%
  mutate(Industry.Group = factor(Industry.Group, Industry.Group)) %>%
  ggplot(aes(x=Total.Equity, y=Free.Cash.Flow, size=Current.Market.Cap, fill=Industry.Group)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    scale_size(range = c(.1, 24), name="Population (M)") +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(legend.position="bottom") +
    ylab("Free.Cash.Flow") +
    xlab("Total.Equity") +
    theme(legend.position = "none")
```

## Bubble chart (Short.and.Long.Term.Debt & Free.Cash.Flow & Total.Equity) 
```{r}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

# Most basic bubble plot
finance_gb %>%
  arrange(desc(Short.and.Long.Term.Debt)) %>%
  mutate(Industry.Group = factor(Industry.Group, Industry.Group)) %>%
  ggplot(aes(x=Free.Cash.Flow, y=Total.Equity, size=Short.and.Long.Term.Debt, fill=Industry.Group)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    scale_size(range = c(.1, 24), name="Population (M)") +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(legend.position="bottom") +
    ylab("Total.Equity") +
    xlab("Free.Cash.Flow") +
    theme(legend.position = "none")
```



## Bubble chart (Current.Market.Cap & ROA & Operating.Return.on.Total.Invested.Capital) 
```{r}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

# Most basic bubble plot
finance_gb %>%
  arrange(desc(Operating.Return.on.Total.Invested.Capital)) %>%
  mutate(Industry.Group = factor(Industry.Group, Industry.Group)) %>%
  ggplot(aes(y=ROA, x=Current.Market.Cap , size=Operating.Return.on.Total.Invested.Capital, fill=Industry.Group)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    scale_size(range = c(.1, 24), name="Population (M)") +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(legend.position="bottom") +
    ylab("ROA") +
    xlab("Current.Market.Cap") +
    theme(legend.position = "none")
```

## Bubble chart (Asset.Turnover & Price.to.Sales.Ratio & ROA) 

```{r}
# # Libraries
# library(ggplot2)
# library(dplyr)
# library(hrbrthemes)
# library(viridis)
# 
# # Most basic bubble plot
# finance_gb %>%
#   arrange(desc(ROA)) %>%
#   mutate(Industry.Group = factor(Industry.Group, Industry.Group)) %>%
#   ggplot(aes(x=Price.to.Sales.Ratio, y=Asset.Turnover , size=ROA, fill=Industry.Group)) +
#     geom_point(alpha=0.5, shape=21, color="black") +
#     scale_size(range = c(.1, 24), name="Population (M)") +
#     scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
#     theme_ipsum() +
#     theme(legend.position="bottom") +
#     ylab("Free.Cash.Flow") +
#     xlab("Total.Equity") +
#     theme(legend.position = "none")

```




-----------------------------------

#  Train-test data set split 
```{r}
finance_df_ml = finance_df[,-1]  # Delete Unix code colounm 'Equity'
finance_df_ml$Industry.Group = as.numeric(finance_df_ml$Industry.Group) 

set.seed(520)
dat.d <- sample(1:nrow(finance_df_ml),size=nrow(finance_df_ml)*0.75,replace = FALSE) #random selection of 75% data.

train <- finance_df_ml[dat.d,] # 75% training data
test <- finance_df_ml[-dat.d,] # remaining 25% test data
```



## Linear Regression 

```{r}
lmFit <- lm(Current.Market.Cap ~., train)

summary(lmFit) # R-square 0.8323
anova(lmFit)
plot(lmFit)



# create the evaluation metrics function

eval_metrics = function(model, df, predictions, target){
    resids = df[,target] - predictions
    resids2 = resids**2
    N = length(predictions)
    r2 = as.character(round(summary(model)$r.squared, 2))
    adj_r2 = as.character(round(summary(model)$adj.r.squared, 2))
    print(adj_r2) #Adjusted R-squared
    print(as.character(round(sqrt(sum(resids2)/N), 2))) #RMSE
}

# # predicting and evaluating the model on train data
# predictions = predict(lmFit, newdata = train)
# eval_metrics(lmFit, train, predictions, target = 'Current.Market.Cap')

# predicting and evaluating the model on test data
predictions = predict(lmFit, newdata = test)
eval_metrics(lmFit, test, predictions, target = 'Current.Market.Cap')

#extractAIC(lmFit2)


library(caret)
data_ctrl <- trainControl(method = "cv", number = 5)
model_caret <- train(Current.Market.Cap ~.,   # model to fit
                     data = train,                        
                     trControl = data_ctrl,              # folds
                     method = "lmStepAIC",                      # specifying regression model
                     na.action = na.pass)

model_caret
summary(model_caret$finalModel)

lmFit2 = lm(Current.Market.Cap ~ Last.Price + Beta + Revenue.Growth.Year.over.Year + 
    ROA + Asset.Turnover + Price.to.Sales.Ratio + Total.Equity + 
    Weighted.Average.Cost.of.Cap + Free.Cash.Flow + Industry.Group + 
    WACC.Economic.Value.Added + Total.Compensation.Paid.to.Executives, data = train)
summary(lmFit2)

# # predicting and evaluating the model on train data
# predictions = predict(lmFit2, newdata = train)
# eval_metrics(lmFit2, train, predictions, target = 'Current.Market.Cap')

# predicting and evaluating the model on test data
predictions = predict(lmFit2, newdata = test)
eval_metrics(lmFit2, test, predictions, target = 'Current.Market.Cap')

```

## LASSO Regression

```{r}
# load packages
library(glmnet)
library(plyr)
library(readr)
library(dplyr)
library(caret)
library(ggplot2)
library(repr)

#define response variable
y_train <- train$Current.Market.Cap
y_test <- test$Current.Market.Cap

#define matrix of predictor variables
x_train <- data.matrix(train[, c("Last.Price","Alpha","Beta",
         "Revenue.Growth.Year.over.Year","ROA","ROE","Profit.Margin",
         "Asset.Turnover","Financial.Leverage","EBIT.Interest","Price.Earnings.Ratio",
         "Price.to.Sales.Ratio","Short.and.Long.Term.Debt","Total.Equity",
         "Operating.Return.on.Total.Invested.Capital","Weighted.Average.Cost.of.Cap",
         "Free.Cash.Flow","Price.to.Book.Ratio","Industry.Group","WACC.Economic.Value.Added",
         "Total.Compensation.Paid.to.Executives")])
x_test <- data.matrix(test[, c("Last.Price","Alpha","Beta",
         "Revenue.Growth.Year.over.Year","ROA","ROE","Profit.Margin",
         "Asset.Turnover","Financial.Leverage","EBIT.Interest","Price.Earnings.Ratio",
         "Price.to.Sales.Ratio","Short.and.Long.Term.Debt","Total.Equity",
         "Operating.Return.on.Total.Invested.Capital","Weighted.Average.Cost.of.Cap",
         "Free.Cash.Flow","Price.to.Book.Ratio","Industry.Group","WACC.Economic.Value.Added",
         "Total.Compensation.Paid.to.Executives")])

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(x_train, y_train, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda   # 1226487108

#produce plot of test MSE by lambda value
plot(cv_model) 

#find coefficients of best model
grid = 10^seq(10,-2,length=100)
plot_model <- glmnet(x_train, y_train, alpha = 1, lambda = grid)
plot(plot_model,"lambda", label=FALSE)

best_model <- glmnet(x_train, y_train, alpha = 1, lambda = best_lambda)
coef(best_model)

#use fitted best model to make predictions
y_predicted <- predict(best_model, s = best_lambda, newx = x_train)

#find SST and SSE
sst <- sum((y_train - mean(y_train))^2)
sse <- sum((y_predicted - y_train)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq  # 0.8240855


# Compute R^2 from true and predicted values

# Prediction and evaluation on test data
predictions_test <- predict(best_model, s = best_lambda, newx = x_test)

#find SST and SSE
sst <- sum((y_test - mean(y_test))^2)
sse <- sum((predictions_test - y_test)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq # -1.551173
```


## Ridge Regression

```{r}
#fit ridge regression model
model <- glmnet(x_train, y_train, alpha = 0)

#view summary of model
summary(model)

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(x_train, y_train, alpha = 0)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda    ## 32576574245

#produce plot of test MSE by lambda value
plot(cv_model) 

#find coefficients of best model
best_model <- glmnet(x_train, y_train, alpha = 0, lambda = best_lambda)
coef(best_model)

#produce Ridge trace plot
plot(model, xvar = "lambda")

#use fitted best model to make predictions
y_predicted <- predict(model, s = best_lambda, newx = x_train)

#find SST and SSE
sst <- sum((y_train - mean(y_train))^2)
sse <- sum((y_predicted - y_train)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq   ## 0.8215536


# Compute R^2 from true and predicted values

# Prediction and evaluation on test data
predictions_test <- predict(best_model, s = best_lambda, newx = x_test)

#find SST and SSE
sst <- sum((y_test - mean(y_test))^2)
sse <- sum((predictions_test - y_test)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq  ## -0.6007401
```



